\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[margin=2.5cm]{geometry}
\usepackage{float}

\title{Multimodal PET/MR Diagnosis with Variational Inference\\Final Report}
\author{BID Lab, School of Biomedical Engineering\\ShanghaiTech University}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We implement a MoPoE-VAE for Alzheimer's Disease diagnosis using multimodal ADNI data. Three synthetic-data experiments compare latent dimension and loss configurations: 256-dim achieves best loss 0.670 (21\% better than 128-dim baseline). On real ADNI data (225 subjects, 11GB), we demonstrate the full pipeline: DCM-NIfTI conversion (77 PET volumes), real-data dataset loading, and end-to-end training with classification evaluation.
\end{abstract}

\section{Dataset}
\textbf{ADNI}: 225 subjects. T1 MRI (223 with NIfTI). PET DICOM converted to NIfTI via SimpleITK (51 FDG + 26 TAU volumes, 6--13 MB each). \textbf{17 subjects} with all three modalities (T1+FDG+TAU) in NIfTI. Total: 11 GB.

\textbf{Diagnosis}: CN (114), EMCI (48), MCI (40), LMCI (21), AD (2).

\section{Method}
MoPoE-VAE: modality-specific 3D CNN encoders, product-of-experts fusion, MLP diagnosis head. Trained with KL divergence + cross-entropy classification loss.

\section{Experiments}

\begin{table}[H]
\centering
\caption{Synthetic data results (100 samples, RTX 4060)}
\begin{tabular}{lcccc}
\toprule
\textbf{Experiment} & \textbf{Latent} & \textbf{Cross} & \textbf{Best Loss} & \textbf{Time} \\
\midrule
Exp1 (Baseline) & 128 & Yes & 0.851 & 220s \\
Exp2 (Large)    & 256 & Yes & \textbf{0.670} & 230s \\
Exp3 (No-Cross) & 128 & No  & 0.763 & 215s \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key finding}: 256-dim latent improves convergence by 21\% over 128-dim baseline.

\begin{table}[H]
\centering
\caption{Real ADNI data configuration}
\begin{tabular}{lr}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Latent dimension & 128 \\
Hidden dimensions & [64, 32] \\
Learning rate & $1\times10^{-3}$ \\
Batch size & 1 (per-subject) \\
Loss & KL + Cross-Entropy \\
Subjects (all-3) & 17 \\
Optimizer & AdamW \\
\bottomrule
\end{tabular}
\end{table}

\section{Infrastructure}
\begin{itemize}
\item \textbf{DCM conversion}: SimpleITK ImageSeriesReader, 77 PET volumes converted
\item \textbf{Data loader}: RealADNIDataset -- reads NIfTI + meta.json, supports modality dropout
\item \textbf{Brain extraction}: HD-BET v2.0.1 ($<$5s per volume)
\item \textbf{Tissue segmentation}: MONAI DynUNet (5.4M params)
\item \textbf{Registration}: SimpleITK (ANTs/FSL replacement)
\item \textbf{Quantization}: bitsandbytes 0.49 (fp16/int8)
\end{itemize}

\section{Pretrained Weights}
No directly applicable pretrained weights exist for the MoPoE-VAE encoder. MONAI segmentation models (DynUNet, SegResNet) can provide encoder initialization through medical image transfer learning. Random Kaiming initialization achieves rapid convergence on both synthetic and real data.

\section{Conclusion}
The MoPoE-VAE architecture with 256-dim latent space achieves the best convergence on synthetic data. The full pipeline -- DCM-to-NIfTI conversion, real-data loading, and training -- is operational on ADNI data. 17 subjects with all three modalities provide a proof-of-concept for cross-modal diagnosis. Future work includes training on the full 38-subject all-3 cohort after DCM conversion completion.

\section{Repository}
\url{https://github.com/tzhazuma/bmebrainproject}

\end{document}
